Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-11-02 submit/apj.tex
more numbers for intrinsic dispersion matrix

The parameterized intrinsic dispersion, $C_c$,  introduces degeneracy in the model, as magnitude and color variation
ascribed to $\Delta$, $\vec{\gamma}^0 k_0$, and $\vec{\gamma}^1 k_1$ can  be attributed to intrinsic dispersion
\color{blue}
and vice versa.
In the case where there is independent intrinsic dispersion in all 5 bands, i.e.\  the input $C_c$ has rank-5, there exists degenerate solutions with a $C_c$ of rank 2.  Those identical predictions generally do not yield the same likelihood, as the determinant of $C_c$ varies over those solutions.  In a maximum-likelihood analysis the rank-2 is preferred over the rank-5 solution, resulting in some intrinsic dispersion getting attributed to $\Delta$, $k_0$, and $k_1$.
In our Bayesian analysis we impose priors on $C_c$, so that a zero-value determinant is disfavored.    
Thus, leakage from per-supernova intrinsic color parameters to intrinsic dispersion could bias the intrinsic color vectors of the SN Ia family,  or conversely, intrinsic dispersion could be spuriously attributed to an intrinsic color.  We examine this effect in the discussion of the results in \S\ref{model2:sec}.
\color{black}
%
%There are several features of the model
%that drive the assignation of variations away from $C_c$:  Maximizing the posterior disfavors the increase of $\det{(C_c)}$
%\color{red}
%and hence favors solutions of $\gamma$'s that minimize the residual dispersion described by $C_c$;
%\color{black}
%The distributions of $\Delta$, $k_0$, and $k_1$ turn out to
%be non-Gaussian, and so are not well described by the Normal covariance we impose in $C_c$; the grey magnitude offsets, $\Delta$, would appear as a constant
%in all elements of the covariance matrix, which is disfavored for the Bayesian prior selected for $C_c$.

@@

Non-trivial residual magnitude
and color dispersions are captured in $C_c$
at the level of 4--10$\sigma$.
The square
root of the diagonal elements are denoted as $\sigma$, which
range from
$\sim 0.01$ to 0.06 mag, significantly smaller
than the dispersion in $\Delta$.
Of particular interest within $C_c$ is intrinsic color dispersion.  The intrinsic dispersions in ${\hat{V}}$ and colors with respect to ${\hat{V}}$ are
\begin{multline}
\sigma({\hat{V}},{\hat{U}}-{\hat{V}},{\hat{B}}-{\hat{V}},{\hat{V}}-{\hat{R}},{\hat{V}}-{\hat{I}})=\\
\left(
0.021^{+0.004}_{-0.005} ,\, 
0.067^{+0.010}_{-0.010},\, 
0.034^{+0.005}_{-0.005},\, 
0.023^{+0.002}_{-0.002},\, 
0.053^{+0.004}_{-0.004}
\right)
 \text{mag}.
 \label{sig_intrinsic:eqn}
 \end{multline}
The correlations between these parameters are
\begin{multline}
Cor({\hat{V}},{\hat{U}}-{\hat{V}},{\hat{B}}-{\hat{V}},{\hat{V}}-{\hat{R}},{\hat{V}}-{\hat{I}})=\\
\begin{pmatrix}
\begin{array}{rrrrr}
\hat{V} & -0.48^{+0.20}_{-0.18} & -0.34^{+0.21}_{-0.22} & 0.88^{+0.10}_{-0.25} & 0.60^{+0.11}_{-0.19} \\
\ldots & {\hat{U}}-{\hat{V}} & 0.60^{+0.10}_{-0.13} & -0.38^{+0.18}_{-0.17} & -0.33^{+0.15}_{-0.14} \\
\ldots  & \ldots & {\hat{B}}-{\hat{V}} & -0.21^{+0.17}_{-0.17} & 0.08^{+0.14}_{-0.15} \\
\ldots  & \ldots  &\ldots  & {\hat{V}}-{\hat{R}} & 0.65^{+0.06}_{-0.07} \\
\ldots  & \ldots  &\ldots  &\ldots & {\hat{V}}-{\hat{I}} \\
\end{array}
 \end{pmatrix}.
 \label{cor_intrinsic:eqn}
 \end{multline}
To reduce clutter only the upper triangular part of the otherwise symmetric matrix is shown.
These color variances are smaller than those derived by \citet{2003A&A...404..901N} or \citet{2007ApJ...659..122J}, for cases where
a direct comparison can be made.
\citetalias{2017ApJ...842...93M}  find an intrinsic dispersion of the SALT2 $c$ parameter (associated with ${\hat{B}}-{\hat{V}}$) of $\approx 0.07$ mag,
though variation of dust properties accounts for some of that dispersion.
%model11_rubin.py
Having no intrinsic color dispersion is disfavored, and not allowing for them can have important consequences.  Running an analysis assuming no intrinsic dispersion (replacing the probabilistic intrinsic magnitudes in Eq.~\ref{ewsiv:eqn}
with a strict equality), we find significant shifts in all slope parameters  and 
artificially
smaller statistical uncertainties.

@@

\color{blue}
We now return to a degeneracy of our models mentioned in \S\ref{modelmodel:sec}.
Parameters are susceptible to the misassignment, in either direction, between intrinsic color, $\Delta$, $k$, and $D$, and intrinsic dispersion $C_c$.
The first two eigenvectors of account for $0.89_{-0.05}^{+0.04}$ of the variance of $C_c$, meaning
that potential leakage of intrinsic colors can account for less than 10\% of the variance, corresponding to 0.024 mag standard deviation.
We establish through simulations that the standard deviations of $C_c$, $\vec{\sigma}$,  are not biased high, which would have been a signature of leakage of
intrinsic color variations from $\delta$ into intrinsic dispersion.
The simulations of input null intrinsic color terms and input intrinsic dispersion do not produce spurious detections of intrinsic color vectors.  We thus conclude that
biases induced by this degeneracy are low relative to our statistical uncertainties.
\color{black}

>>>>

The parameterized intrinsic dispersion, $C_c$,  introduces degeneracy in the model, as magnitude and color variation
ascribed to $\Delta$, $\vec{\gamma}^0 k_0$, and $\vec{\gamma}^1 k_1$ can  be attributed to intrinsic dispersion
\color{blue}
and vice versa.
In the case where there is independent intrinsic dispersion in all 5 bands, i.e.\  the input $C_c$ has rank-5, there exists degenerate solutions with a $C_c$ of rank 2.  Those identical predictions generally do not yield the same likelihood, as the determinant of $C_c$ varies over those solutions.  In a maximum-likelihood analysis the rank-2 is preferred over the rank-5 solution, resulting in some intrinsic dispersion getting attributed to $\Delta$, $k_0$, and $k_1$.
In our Bayesian analysis the priors, though not totally disallowing rank-2 $C_c$, favor rank-5  and prevent convergence to this solution.    
Leakage from per-supernova intrinsic color parameters to intrinsic dispersion could bias the intrinsic color vectors of the SN Ia family,  or conversely, intrinsic dispersion could be spuriously attributed to an intrinsic color.  We examine this effect in the discussion of the results in \S\ref{model2:sec}.
\color{black}
%
%There are several features of the model
%that drive the assignation of variations away from $C_c$:  Maximizing the posterior disfavors the increase of $\det{(C_c)}$
%\color{red}
%and hence favors solutions of $\gamma$'s that minimize the residual dispersion described by $C_c$;
%\color{black}
%The distributions of $\Delta$, $k_0$, and $k_1$ turn out to
%be non-Gaussian, and so are not well described by the Normal covariance we impose in $C_c$; the grey magnitude offsets, $\Delta$, would appear as a constant
%in all elements of the covariance matrix, which is disfavored for the Bayesian prior selected for $C_c$.

@@

Non-trivial residual magnitude
and color dispersions are captured in $C_c$
at the level of 4--10$\sigma$.
The square
root of the diagonal elements are denoted as $\sigma$, which
range from
$\sim 0.01$ to 0.06 mag, significantly smaller
than the dispersion in $\Delta$.
\color{green}
The posterior of $\sigma_{\hat{R}}$ seen in Figure~\ref{global4:fig} does not exclude values that approach zero, highlighting that non-rank 5
$C_c$'s are accommodated by our prior.
\color{black}
Of particular interest within $C_c$ is intrinsic color dispersion.  The intrinsic dispersions in ${\hat{V}}$ and colors with respect to ${\hat{V}}$ are
\begin{multline}
\sigma({\hat{V}},{\hat{U}}-{\hat{V}},{\hat{B}}-{\hat{V}},{\hat{V}}-{\hat{R}},{\hat{V}}-{\hat{I}})=\\
\left(
0.021^{+0.004}_{-0.005} ,\, 
0.067^{+0.010}_{-0.010},\, 
0.034^{+0.005}_{-0.005},\, 
0.023^{+0.002}_{-0.002},\, 
0.053^{+0.004}_{-0.004}
\right)
 \text{mag}.
 \label{sig_intrinsic:eqn}
 \end{multline}
The correlations between these parameters are
\begin{multline}
Cor({\hat{V}},{\hat{U}}-{\hat{V}},{\hat{B}}-{\hat{V}},{\hat{V}}-{\hat{R}},{\hat{V}}-{\hat{I}})=\\
\begin{pmatrix}
\begin{array}{rrrrr}
\hat{V} & -0.48^{+0.20}_{-0.18} & -0.34^{+0.21}_{-0.22} & 0.88^{+0.10}_{-0.25} & 0.60^{+0.11}_{-0.19} \\
\ldots & {\hat{U}}-{\hat{V}} & 0.60^{+0.10}_{-0.13} & -0.38^{+0.18}_{-0.17} & -0.33^{+0.15}_{-0.14} \\
\ldots  & \ldots & {\hat{B}}-{\hat{V}} & -0.21^{+0.17}_{-0.17} & 0.08^{+0.14}_{-0.15} \\
\ldots  & \ldots  &\ldots  & {\hat{V}}-{\hat{R}} & 0.65^{+0.06}_{-0.07} \\
\ldots  & \ldots  &\ldots  &\ldots & {\hat{V}}-{\hat{I}} \\
\end{array}
 \end{pmatrix}.
 \label{cor_intrinsic:eqn}
 \end{multline}
To reduce clutter only the upper triangular part of the otherwise symmetric matrix is shown.
These color variances are smaller than those derived by \citet{2003A&A...404..901N} or \citet{2007ApJ...659..122J}, for cases where
a direct comparison can be made.
\citetalias{2017ApJ...842...93M}  find an intrinsic dispersion of the SALT2 $c$ parameter (associated with ${\hat{B}}-{\hat{V}}$) of $\approx 0.07$ mag,
though variation of dust properties accounts for some of that dispersion.
%model11_rubin.py
Having no intrinsic color dispersion is disfavored, and not allowing for them can have important consequences.  Running an analysis assuming no intrinsic dispersion (replacing the probabilistic intrinsic magnitudes in Eq.~\ref{ewsiv:eqn}
with a strict equality), we find significant shifts in all slope parameters  and 
artificially
smaller statistical uncertainties.

@@

\color{blue}
We now return to a degeneracy of our models mentioned in \S\ref{modelmodel:sec}.
Parameters are susceptible to the misassignment, in either direction, between intrinsic color, $\Delta$, $k$, and $D$, and intrinsic dispersion $C_c$.
The first two eigenvectors of $C_c$ account for $0.89_{-0.05}^{+0.04}$ of its variance  meaning
that potential leakage of intrinsic colors can account for less than 10\% of the variance, corresponding to 0.024 mag standard deviation.
We establish through simulations that the standard deviations of $C_c$, $\vec{\sigma}$,  are not biased high, which would have been a signature of leakage of
intrinsic color variations from $\delta$ into intrinsic dispersion.
The simulations of input null intrinsic color terms and input intrinsic dispersion do not produce spurious detections of intrinsic color vectors.  We thus conclude that
biases induced by this degeneracy are low relative to our statistical uncertainties.
\color{black}

\color{green}
Contour of sigma and delta.

Results from diagonal covariance.
\color{black}
